---
title: "SIA exploratory"
format: gfm
editor: visual
---

## Load packages and save formatting preferences

```{r}
#| include: false
library(patchwork)
library(broom)
library(tidyverse)
library(GGally)
library(maps)
library(rnaturalearth)
library(sf)
library(gt)


my_theme <- theme_minimal() +
  theme(
    axis.title.x = element_text(size = 30, margin = margin(t = 20)),
    axis.title.y = element_text(size = 30, margin = margin(r = 20)),
    axis.text.x = element_text(size = 24),
    axis.text.y = element_text(size = 24),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    legend.title = element_text(size = 30, face = "bold"),
    legend.text = element_text(size = 24),
    strip.text = element_text(size = 30, face = "bold"),
    axis.line = element_line(colour = "black", linewidth = 0.8)) 

cbf <- c("orange", "#56B4E9", "#009E73", "#F0E442","pink")

```

## Input data and merge into one

```{r}
#| echo: true
lab_outputs<-read.csv("Lab_outputs.csv",header=TRUE)
head(lab_outputs)
lab_outputs <- lab_outputs[, 1:(ncol(lab_outputs) - 4)]

metadata<-read.csv("Metadata.csv",header=TRUE)
metadata <- metadata %>%
  mutate(Turtle_ID = row_number())

lab_outputs <- lab_outputs %>%
  mutate(Sample_ID = case_when(
    Sample_type == "Blood" ~ sub("^RBC([0-9]+)[a-zA-Z]$", "\\1", Sample_name), 
    Sample_type == "Skin"  ~ Sample_name,                       
    TRUE ~ Sample_name
  ))
head(lab_outputs)

blood_labs <- lab_outputs %>%
  filter(Sample_type == "Blood") %>%
  mutate(Sample_ID = as.integer(Sample_ID)) %>%
  left_join(metadata %>% select(Blood_RB, Sex, Life_stage, Date, Mass, Adipose_perc,Turtle_ID, Latitude, Longitude),
            by = c("Sample_ID" = "Blood_RB")) %>%
  mutate(Sample_ID = as.character(Sample_ID))

skin_labs <- lab_outputs %>%
  filter(Sample_type == "Skin") %>%
  left_join(metadata %>% select(Biopsy_no, Genetics_no, Sex, Life_stage, Date, Mass,Adipose_perc,Turtle_ID, Latitude, Longitude) %>%
              pivot_longer(c(Biopsy_no, Genetics_no), values_to = "Sample_ID") %>%
              select(-name),
            by = "Sample_ID")

SIA_full <- bind_rows(blood_labs, skin_labs) %>%
  select(-Sample_ID)

SIA_full <- SIA_full %>%
  mutate(
    Month = month(Date),
    Season = if_else(Month %in% 11:12 | Month %in% 1:4, "Wet", "Dry")
  )
SIA_full$Date <- as.Date(SIA_full$Date, format = "%d/%m/%Y")
SIA_full$Year <- year(SIA_full$Date)
SIA_full$Year <- as.factor(SIA_full$Year)
SIA_full <- SIA_full %>%
  mutate(Mass = as.numeric(Mass))

head(SIA_full)
```

## Summmary table by sample type

```{r}
SIA_full %>%
  group_by(Sample_type, Sex, Season) %>%
  summarise(
    mean_C = round(mean(C, na.rm = TRUE), 2),
    sd_C   = round(sd(C, na.rm = TRUE), 2),
    mean_N = round(mean(N, na.rm = TRUE), 2),
    sd_N   = round(sd(N, na.rm = TRUE), 2),
    n      = n(),
    .groups = "drop"
  )
```

## When were samples taken?

```{r}

SIA_full <- SIA_full %>%
  mutate(Month_Year = format(Date, "%m-%y")) %>%
  mutate(Month_Year = factor(Month_Year, levels = unique(Month_Year[order(Date)])))
SIA_full %>%
  ggplot(aes(x = Month_Year)) +
  geom_bar(fill = "#4A90C4") +
  labs(
    x = "Trip (Month-Year)",
    y = "Number of samples"
  ) +
  my_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Where were samples taken?

```{r}
SIA_full$Longitude <- as.numeric(as.character(SIA_full$Longitude))
SIA_full$Location <- ifelse(SIA_full$Longitude < 122.07, "EP", "RB")

world <- ne_countries(scale = "medium", returnclass = "sf")

geo.box <- c(xmin= 110, xmax= 140, ymin= -23, ymax= -8)
sf::sf_use_s2(FALSE)
aus_map <- sf::read_sf("~/Documents/Murdoch/Data/Chapter 1/Additional_tracks/Final/gshhg-shp-2.3.7/GSHHS_shp/f/GSHHS_f_L1.shp") %>% st_crop(geo.box)

ggplot() +
  geom_sf(data = aus_map, fill = "grey90", color = "grey40") +
  geom_point(data = SIA_full, aes(x = Longitude, y = Latitude, color = Location),
             size = 2) +
  coord_sf(xlim = c(121.9, 122.4),
           ylim = c(-18.3, -17.8)) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude", color = "Location")

SIA_full$Location<-as.factor(SIA_full$Location)

```

## Exploring inter-annual trends in biopsy data

```{r}
#| echo: true

SIA_biopsy<-SIA_full[SIA_full$Sample_type=="Skin",]
```

## Plot for SIA

```{r}
SIA_biopsy %>%
  select(C, N, Mass, Season) %>%
  ggpairs()
```

```{r}
SIA_biopsy %>%
  group_by(Year, Sex) %>%
  summarise(mean_C = mean(C, na.rm = TRUE),
            se_C = sd(C, na.rm = TRUE) / sqrt(n()),
            n = n(),
            .groups = "drop") %>%
  ggplot(aes(x = Year, y = mean_C, colour = Sex, group = Sex)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_C - se_C, ymax = mean_C + se_C), width = 0.3, alpha = 0.6)+
  scale_colour_manual(values = c("#4A90C4", "darkgreen")) +
  labs(
    x = "Year",
    y = "Mean dC",
    colour = "Sex"
  ) +
  my_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

SIA_biopsy %>%
  group_by(Year, Sex) %>%
  summarise(mean_N = mean(N, na.rm = TRUE),
            se_N = sd(N, na.rm = TRUE) / sqrt(n()),
            n = n(),
            .groups = "drop") %>%
  ggplot(aes(x = Year, y = mean_N, colour = Sex, group = Sex)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_N - se_N, ymax = mean_N + se_N), width = 0.3, alpha = 0.6)+
  scale_colour_manual(values = c("#4A90C4", "darkgreen")) +  
  labs(
    x = "Year",
    y = "Mean dN",
    colour = "Sex"
  ) +
  my_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Isoscape plots by year

```{r}
#| echo: true
SIA_biopsy %>%
  ggplot(aes(x = C, y = N, colour = Season, shape = Sex)) +
  geom_point(size = 3, alpha = 0.7) +
  facet_wrap(~Year)

first.plot <- ggplot(data = SIA_biopsy, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Year, shape = Year), size = 5) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_biopsy %>% 
  group_by(Year) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Year), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Year, 
                   fill = Year, 
                   color = Year), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by sex

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_biopsy, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Sex, shape = Sex), size = 5) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_biopsy %>% 
  group_by(Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Sex), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Sex, 
                   fill = Sex, 
                   color = Sex), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by year + sex

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_biopsy, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Year), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_biopsy %>% 
  group_by(Year, Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Year), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Year, 
                   fill = Year, 
                   color = Year), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by year + sex + season

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_biopsy, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Year, shape = Season), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_biopsy %>% 
  group_by(Year, Sex, Season) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Year, shape = Season), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =interaction(Year, Season), 
                   fill = Year, 
                   color = Year,
                   linetype = Season), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Running models

```{r}
#| echo: true
SIA_biopsy %>% count(Year, Sex, Season)

mC <- lm(C ~ Year + Sex + Season, data = SIA_biopsy)
summary(mC)

shapiro.test(resid(mC))

par(mfrow = c(2,2))
plot(mC)

mN <- lm(N ~ Year + Sex + Season, data = SIA_biopsy)
summary(mN)

shapiro.test(resid(mN))

plot(mN)

mC <- lm(C ~ Year + Sex, data = SIA_biopsy)
summary(mC)

shapiro.test(resid(mC))

par(mfrow = c(2,2))
plot(mC)

mN <- lm(N ~ Year + Sex, data = SIA_biopsy)
summary(mN)

shapiro.test(resid(mN))

plot(mN)
```

```{r}
#| echo: true

SIA_biopsy %>%
  ggplot(aes(x = C, y = Mass)) +
  geom_point(size = 5)

SIA_biopsy %>%
  ggplot(aes(x = N, y = Mass)) +
  geom_point(size = 5)
```

## Next on the list is repeating this for blood samples AND tissue comparison AND considering the above analysis and whether the sample is actually representative of wet/dry

##Consider not using season since turnover time is likely longer than a season BUT if we can validate against eDNA we may be able to infer expected turnover time

##Blood turnover expected 3-5 months and skin could be 3-12 months (Seminoff et al., 2007)

## Exploring inter-annual trends in biopsy data

```{r}
#| echo: true

SIA_blood<-SIA_full[SIA_full$Sample_type=="Blood",]
```

## Plot for SIA

```{r}
SIA_blood %>%
  select(C, N, Mass, Season) %>%
  ggpairs()
```

```{r}

SIA_blood %>%
  group_by(Month_Year, Sex) %>%
  summarise(mean_C = mean(C, na.rm = TRUE),
            se_C = sd(C, na.rm = TRUE) / sqrt(n()),
            n = n(),
            .groups = "drop") %>%
  ggplot(aes(x = Month_Year, y = mean_C, colour = Sex, group = Sex)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_C - se_C, ymax = mean_C + se_C), width = 0.3, alpha = 0.6)+
  scale_colour_manual(values = c("#4A90C4", "darkgreen")) +
  labs(
    x = "Month-Year",
    y = "Mean dC",
    colour = "Sex"
  ) +
  my_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

SIA_blood %>%
  group_by(Month_Year, Sex) %>%
  summarise(mean_N = mean(N, na.rm = TRUE),
            se_N = sd(N, na.rm = TRUE) / sqrt(n()),
            n = n(),
            .groups = "drop") %>%
  ggplot(aes(x = Month_Year, y = mean_N, colour = Sex, group = Sex)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = mean_N - se_N, ymax = mean_N + se_N), width = 0.3, alpha = 0.6)+
  scale_colour_manual(values = c("#4A90C4", "darkgreen")) +  
  labs(
    x = "Month-Year",
    y = "Mean dN",
    colour = "Sex"
  ) +
  my_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Isoscape plots by season

```{r}
#| echo: true
SIA_blood %>%
  ggplot(aes(x = C, y = N, colour = Season, shape = Season)) +
  geom_point(size = 3, alpha = 0.7) +
  facet_wrap(~Season)

first.plot <- ggplot(data = SIA_blood, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Season, shape = Season), size = 5) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_blood %>% 
  group_by(Season) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Season), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Season, 
                   fill = Season, 
                   color = Season), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by sex

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_blood, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Sex, shape = Sex), size = 5) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_blood %>% 
  group_by(Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Sex), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Sex, 
                   fill = Sex, 
                   color = Sex), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by year + sex

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_blood, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Season), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_blood %>% 
  group_by(Year, Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Year), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Year, 
                   fill = Year, 
                   color = Year), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

## Isoscape plots by year + sex + season

```{r}
#| echo: true
first.plot <- ggplot(data = SIA_blood, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Month_Year), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_blood %>% 
  group_by(Month_Year, Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Month_Year), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group = Month_Year, 
                   fill = Month_Year, 
                   color = Month_Year,
                   linetype = Month_Year), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)
```

##Try inter-annual trends again with just August/dry season data OR include month as a covariate, especially since season was significant for N

```{r}

```


## Running models

```{r}
#| echo: true
SIA_blood %>% count(Year, Sex, Season)

mC <- lm(C ~ Year + Sex + Season, data = SIA_blood)
summary(mC)

shapiro.test(resid(mC))

par(mfrow = c(2,2))
plot(mC)

mN <- lm(N ~ Year + Sex + Season, data = SIA_blood)
summary(mN)

shapiro.test(resid(mN))

plot(mN)

```

```{r}
#| echo: true

SIA_blood %>%
  ggplot(aes(x = C, y = Mass)) +
  geom_point(size = 5)

SIA_blood %>%
  ggplot(aes(x = N, y = Mass)) +
  geom_point(size = 5)

summary(lm(C ~ Mass + Season, data = SIA_blood))
summary(lm(N ~ Mass +  Season, data = SIA_blood))

#Significant effect of mass for nitrogen, both significant when you take out sex

SIA_blood %>%
  ggplot(aes(x = C, y = Adipose_perc)) +
  geom_point(size = 5)

SIA_blood %>%
  ggplot(aes(x = N, y = Adipose_perc)) +
  geom_point(size = 5)

summary(lm(C ~ Adipose_perc + Sex, data = SIA_blood))
summary(lm(N ~ Adipose_perc +  Sex, data = SIA_blood))
```

## Can also try with reproductive state, once validated and tidied up

##Tissue comparison for those with both biopsy and bloods

```{r}
paired_ids <- SIA_full %>%
  group_by(Turtle_ID) %>%
  summarise(has_skin = any(Sample_type == "Skin"),
            has_blood = any(Sample_type == "Blood")) %>%
  filter(has_skin & has_blood) %>%
  pull(Turtle_ID)

SIA_both <- SIA_full %>%
  filter(Turtle_ID %in% paired_ids)

n_distinct(SIA_both$Turtle_ID)

head(SIA_both)

#TDFS

discrim <- tribble(
  ~Sample_type, ~Isotope, ~TDF, ~TDF_sd,
  "Blood", "C", 0.46, 0.35,
  "Blood", "N", 1.49, 0.76,
  "Skin",  "C", 2.26, 0.61,
  "Skin",  "N", 1.85, 0.50
)

SIA_both_corrected <- SIA_both %>%
  left_join(discrim %>% filter(Isotope == "C") %>% select(Sample_type, TDF_C = TDF), by = "Sample_type") %>%
  left_join(discrim %>% filter(Isotope == "N") %>% select(Sample_type, TDF_N = TDF), by = "Sample_type") %>%
  mutate(C_diet = C - TDF_C,
         N_diet = N - TDF_N)

SIA_both_long <- SIA_both_corrected %>%
  select(Turtle_ID, Sample_type, C, N) %>%
  pivot_longer(cols = c(C, N), names_to = "Isotope", values_to = "Value") %>%
  mutate(Sample_type = factor(Sample_type, levels = c("Blood", "Skin")))

ggplot(SIA_both_long, aes(x = Sample_type, y = Value, group = Turtle_ID)) +
  geom_line(alpha = 0.3, colour = "grey40") +
  geom_point(aes(colour = Sample_type), size = 4, alpha = 0.8) +
  facet_wrap(~Isotope, scales = "free_y", 
             labeller = as_labeller(c(C = "δ13C (‰)", N = "δ15N (‰)"))) +
  scale_colour_manual(values = c("#4A90C4", "darkgreen")) +
  labs(x = "Tissue", y = NULL, colour = "Tissue") +
  my_theme


#48 individuals with both biopsy and bloods

first.plot <- ggplot(data = SIA_both_corrected, 
                   mapping = aes(x = C_diet, y = N_diet)) + 
  geom_point(aes(color = Sample_type), size = 5) +
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size = 20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size = 35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_manual(values = c("#4A90C4", "darkgreen"))
print(classic.first.plot)

fbmeans <- SIA_both_corrected %>% 
  group_by(Sample_type) %>% 
  summarise(count = n(),
            mC = mean(C_diet), sdC = sd(C_diet), 
            mN = mean(N_diet), sdN = sd(N_diet))
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0) +
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, y = mN, fill = Sample_type), 
             color = "black", shape = 22, size = 6,
             alpha = 0.9, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group = Sample_type, 
                   fill = Sample_type, 
                   color = Sample_type), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon") +
  coord_equal()
print(ellipse.plot3)
```

##Within-individual differences??

```{r}
individual_shift <- SIA_both_corrected %>%
  select(Turtle_ID, Sample_type, C_diet, N_diet) %>%
  pivot_wider(names_from = Sample_type, values_from = c(C_diet, N_diet)) %>%
  mutate(
    C_shift = C_diet_Skin - C_diet_Blood,   # positive = long-term diet more 13C-enriched than recent
    N_shift = N_diet_Skin - N_diet_Blood    # positive = long-term diet higher trophic level than recent
  )

individual_shift %>%
  ggplot(aes(x = C_shift)) +
  geom_histogram(bins = 15, fill = "#4A90C4", colour = "white") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey30") +
  labs(x = "Skin − Blood δ13C shift (diet-corrected, ‰)", y = "Number of individuals") +
  my_theme

shift_model <- lm(C_shift ~ Sex + Year, data = individual_shift %>% left_join(SIA_both_corrected %>% distinct(Turtle_ID, Sex, Year), by = "Turtle_ID"))
summary(shift_model)
```

##Do we have any individuals with repeat bloods?

```{r}

#requires DBCA data from export

```

##Explore influence of rainfall or SST

## Explore spatial differences in isotopes. Looks like there is signal.

#Interestingly looking at satellite tracks individuals appear to go to entrance point first after breeding, and by march/april end up back in the bay

```{r}

first.plot <- ggplot(data = SIA_blood, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Location), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_blood %>% 
  group_by(Location, Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Location), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Location, 
                   fill = Location, 
                   color = Location), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)

summary(lm(C ~ Location + Sex + Season, data = SIA_blood))

first.plot <- ggplot(data = SIA_biopsy, 
                   mapping = aes(x = C, 
                                 y = N)) + 
  geom_point(aes(color = Location), size = 5) +
  facet_wrap(~Sex)+
  ylab(expression(paste(delta^{15}, "N (\u2030)"))) +
  xlab(expression(paste(delta^{13}, "C (\u2030)"))) + 
  theme(text = element_text(size=20))

classic.first.plot <- first.plot + theme_classic() + 
  theme(text = element_text(size=35)) + 
  coord_equal() + 
  theme(axis.ticks.length = unit(-0.2, "cm")) +
  scale_colour_viridis_d(end = 0.9)
print(classic.first.plot)

fbmeans <- SIA_biopsy %>% 
  group_by(Location, Sex) %>% 
  summarise(count = n(),
            mC = mean(C), 
            sdC = sd(C), 
            mN = mean(N), 
            sdN = sd(N) )
print(fbmeans)

second.plot <- first.plot + 
  geom_errorbar(data = fbmeans, 
                mapping = aes(x = mC, y = mN,
                              ymin = mN - 1.96*sdN, 
                              ymax = mN + 1.96*sdN), 
                width = 0)+
  geom_errorbarh(data = fbmeans, 
                 mapping = aes(x = mC, y = mN,
                               xmin = mC - 1.96*sdC,
                               xmax = mC + 1.96*sdC),
                 height = 0) + 
  geom_point(data = fbmeans, aes(x = mC, 
                                 y = mN,
                                 fill = Location), 
             color = "black", shape = 22, size = 5,
             alpha = 0.7, show.legend = FALSE) + 
  coord_equal()
print(second.plot)

p.ell <- 0.95

ellipse.plot3 <- first.plot + 
  stat_ellipse(aes(group =Location, 
                   fill = Location, 
                   color = Location), 
               alpha = 0.25, 
               level = p.ell,
               type = "t",
               geom = "polygon")
print(ellipse.plot3)

summary(lm(C ~ Location, data = SIA_biopsy))
summary(lm(C ~ Location + Season + Sex, data = SIA_biopsy))
```

##Sample size tables for different analyses

```{r}
sample_table <- SIA_biopsy %>%
  count(Sex, Year, Season)

wide_data <- sample_table %>%
  unite("Year_Season", Year, Season, sep = "_") %>%
  pivot_wider(names_from = Year_Season, values_from = n, values_fill = 0)

col_years <- sample_table %>% 
  distinct(Year, Season) %>% 
  arrange(Year, Season) %>%
  unite("Year_Season", Year, Season, sep = "_", remove = FALSE)

wide_data <- wide_data %>%
  select(Sex, all_of(col_years$Year_Season))

gt_table <- wide_data %>%
  gt(rowname_col = "Sex")

gt_table <- wide_data %>%
  gt(rowname_col = "Sex") %>%
  cols_label(.list = setNames(
    as.list(sub("^[0-9]+_", "", col_years$Year_Season)), 
    col_years$Year_Season
  ))

# add a spanner per year
for (yr in unique(col_years$Year)) {
  yr_cols <- col_years$Year_Season[col_years$Year == yr]
  gt_table <- gt_table %>%
    tab_spanner(label = yr, columns = all_of(yr_cols))
}


gt_table
```

```{r}
sample_table <- SIA_blood %>%
  count(Sex, Month_Year)

wide_data <- sample_table %>%
  pivot_wider(names_from = Month_Year, values_from = n, values_fill = 0)

wide_data %>%
  gt(rowname_col = "Sex") %>%
  cols_label(Sex = "Sex")
```

```{r}

SIA_full_EP <- SIA_full[which(SIA_full$Location == "EP"), ]

sample_table <- SIA_full_EP %>%
  count(Sex, Sample_type)

wide_data <- sample_table %>%
  pivot_wider(names_from = Sample_type, values_from = n, values_fill = 0)

wide_data %>%
  gt(rowname_col = "Sex") %>%
  cols_label(Sex = "Sex")
```
```

